---
title: 'Marcell Experimental Forest: S2 Stripchart Stream Height (2017 - 2021)'
author: "Mia Forsline"
date: '`r format(Sys.time(), "%B %d, %Y")`'
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, output_file="/Users/miaforsline/Desktop/water_years/index.html") })
output:
  html_document:
    theme: flatly
    code_folding: hide
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 72
---

# Objectives

The goal of this step in the workflow is to examine the stripchart
streamflow data for the S2 Marcell Experimental Forest site for the
water years of 2017 - 2021. This RMD is the first in the workflow and
reads in, cleans, and visualizes the stripchart stream height data for
the 2017 - 2021 water years at the S2 bog site of the [Marcell
Experimental Forest, Northern Research
Station](https://www.nrs.fs.fed.us/ef/marcell/).

# Set Up

-   Install/load necessary R packages
-   Set working directory if necessary (or create a file path to use
    throughout the RMD to call the data from Box)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE, 
                      warning = FALSE)

options(knitr.duplicate.label = "allow")

#install packages if necessary, then load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(
  here, #for file paths 
  janitor, 
  plotly, #for creating interactive plots 
  readxl, 
  render, 
  tidyverse,
  zoo #for interpolating missing data points 
  )

# Set working directory path
## Nina's working directory path 
#setwd("~/../../Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts")
```

# Water Year 2017

-   Read in the WY 2017 data from the External-MEF_DATA Box folder
    -   Specifically, raw data for the S2 Bog site is found at the
        following file path:
        `External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2`
-   Clean and manipulate the WY 2017 data into a tidy format
    -   Rename columns to use lower_snake_case and avoid special
        characters
    -   Create a new `date` column that follows the format YYYY-MM-DD
        (and does not include a timestamp) for ease of
        sorting/manipulating dates
-   Note that this cleaning process will repeat for each individual
    water year

```{r wy2017}
#create file path to call the data from Box 
## Mia's file path 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#read in the 2017 data 
wy2017 <- read_excel(path = here(filepath, "WY2017.S2_Breakpoint.xlsx"), 
                     skip =3) #skip the top 3 rows of the Excel Sheet 

#clean the 2017 data
wy2017_clean <- wy2017 %>% 
  #change column names
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  #remove unnecessary columns 
  select(c("datetime", "stream_height_ft")) %>% 
  #format column names into lower_snake_case
  clean_names() %>% 
  #create a date column 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT"))
```

-   Test that the manipulated data frame has the same number of rows as
    the original data frame

    -   We do not want to accidentally lose data, so the code will throw
        an error is the two data frames do not match

```{r wy2017_test}
#test the 2017 data
if(nrow(wy2017) != nrow(wy2017_clean)) stop("Check clean dataframe dimensions")
```

-   Plot the WY 2017 data

```{r wy2017_plot}
#sample size
n <- nrow(wy2017_clean)

#plot 
p_2017 <- ggplot(data = wy2017_clean) + 
  geom_line(aes(x = datetime, 
                y = stream_height_ft,
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3)))) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Stream Height (ft)", 
       title = "WY 2017 S2 Bog Weir Stream Height", 
       caption = paste0("n = ", n)) + 
  theme(plot.title = element_text(hjust = 0.5))

#static plot 
#p_2017

#interactive plot 
ggplotly(p_2017, 
         tooltip = "text") %>% 
  layout(
    title = list(text = paste0("WY 2017 S2 Bog Weir Stream Height",
                                    '<br>',
                                    '<sup>',
                                     #subtitle
                                     paste0("n = ", n), 
                                    '</sup>')), 
         #title size 
         font = list(size = 14))
```

# Water Year 2018

-   Read in the WY 2018 data from the External-MEF_DATA Box folder
    -   Raw data for the S2 Bog site is found at the following file
        path:
        `External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2`
-   Clean and manipulate the WY 2018 data into a tidy format
-   Change stream height values of 0 to missing NA values during the
    following time periods:
    -   *December 1-3 clock stopped, no record*

    -   *December 8-10 pin reached end of chart, no record*

    -   *Dec 11-end, clock is barely running. No way to tell which days
        are what on the chart.*

    -   *November 10-14 clock stopped, no record*

    -   *November 25- December 4 clock stopped, no record Possibly the
        clock was moving super slowly for some of it?*

    -   *For Oct 9-10 there was no rain for flow to increase*

    -   *Oct 11-15 clock was stuck and the pin did not move forward in
        time*
-   NA values are derived from notes recorded in .txt files found at the
    following Box filepath:
    `External-MEF_DATA/Hydro/Streamflow/L0_daily/StripCharts/DailyBreakpoint/2018`

```{r wy2018}
#create file path to call the data from Box 
## Mia's file path 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#create a function to specify custom time intervals 
`%between%` <- function(x, interval) x >= interval[1] & x <= interval[2]

#read in the 2018 data 
wy2018 <- read_excel(path = here(filepath, "WY2018.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2018 data
wy2018_clean <- wy2018 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#assign October NA values 
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-10-09', '2018-10-10'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-10-11', '2018-10-15'))] <- NA

#assign November NA values 
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-11-10', '2018-11-14'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-11-25', '2018-12-04'))] <- NA

#assign December NA values 
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-12-01', '2018-12-03'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-12-08', '2018-12-10'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-12-11', '2018-12-31'))] <- NA
```

-   Test that the manipulated data frame has the same number of rows as
    the original data frame

```{r wy2018_test}
#test the 2018 data
if(nrow(wy2018) != nrow(wy2018_clean)) stop("Check clean dataframe dimensions")
```

-   Plot the WY 2018 data

```{r wy2018_plot}
#sample size 
n <- nrow(wy2018_clean)

#plot
p_2018_clean <- ggplot(data = wy2018_clean) + 
  geom_line(aes(x = datetime, 
                y = stream_height_ft,
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3)))) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Stream Height (ft)", 
       title = "WY 2018 S2 Bog Weir Stream Height (Unaltered)") + 
  theme(plot.title = element_text(hjust = 0.5))

#note the weird oscillating values from Oct 31 - Nov 1
#static plot
#p_2018_clean

#interactive plot 
ggplotly(p_2018_clean, tooltip = "text") %>% 
  layout(
    title = list(text = paste0("WY 2018 S2 Bog Weir Stream Height",
                                    '<br>',
                                    '<sup>',
                                     #subtitle
                                     paste0("n = ", n), 
                                    '</sup>')), 
         #title size 
         font = list(size = 14))
```

-   Next, we need to remove unusual stream height values from Oct 31 -
    Nov 1, 2018. During this time period, the stream height values
    oscillated rapidly, likely indicating human error in recording the
    data. Thus, we need to remove the erroneous values to smooth out the
    line. For more details see the next code chunk below.

```{r wy2018_altered}
wy2018_altered <- wy2018_clean

#for now, remove erroneous values < 0.10 during the time period of interest 
wy2018_altered <- subset(wy2018_altered, !(wy2018_altered$date %between% as.Date(c('2018-10-31', '2018-11-01')) & wy2018_altered$stream_height_ft < 0.10))

#test the 2018 altered data - there should be fewer rows in the altered dataframe than the original/cleaned dataframe 
if(nrow(wy2018_altered) == nrow(wy2018_clean)) stop("Check altered dataframe dimensions")
```

-   A closer look at the wonky oscillating values and how the data looks
    after the values are removed

```{r wy2018_altered_plot}
#sample size 
n <- nrow(wy2018_altered)

#plot
p_2018_altered <- ggplot(data = wy2018_altered) + 
  geom_line(aes(x = datetime, 
                y = stream_height_ft, 
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3)))) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Stream Height (ft)", 
       title = "WY 2018 S2 Bog Stream Height (Altered)") + 
  theme(plot.title = element_text(hjust = 0.5))

#note that the wonky values have been removed and the graphis now smoothed out 
#static plot
#p_2018_altered

#interactive plot 
ggplotly(p_2018_altered, tooltip = "text") %>% 
  layout(
    title = list(text = paste0("WY 2018 S2 Bog Weir Stream Height",
                                    '<br>',
                                    '<sup>',
                                     #subtitle
                                     paste0("n = ", n), 
                                    '</sup>')), 
         #title size 
         font = list(size = 14))
```

## Altered Oct 31 - Nov 1, 2018 Streamflow Data

During this time, it is most likely that the dates were read/input
incorrectly, resulting in an unnatural oscillation of stream flow values
on Oct 31 and Nov 1 of 2018. Therefore, we removed values \< 0.10 during
this time period to smooth out the overall curve of stream height
values.

```{r oct31-nov1}
#subset the time period of interest
subset_2018 <- wy2018_clean %>% 
  subset(date %between% as.Date(c('2018-10-30', '2018-11-02'))) 

#plot the data 
p_subset <- ggplot(data = subset_2018) + 
  geom_line(aes(x = datetime, y = stream_height_ft)) + 
  labs(title = "Wonky Oct 31 - Nov 1 Streamflow Values")

#static plot 
p_subset

#interactive plot
#ggplotly(p_subset)

#remove unusual values below 0.10 
subset_2018_altered <- wy2018_clean %>% 
  subset(date %between% as.Date(c('2018-10-30', '2018-11-02'))) %>% 
  subset(stream_height_ft > 0.10)

#plot the altered data
p_subset_altered <- ggplot(data = subset_2018_altered) + 
  geom_line(aes(x = datetime, y = stream_height_ft)) + 
  labs(title = "Fixed Oct 31 - Nov 1 Streamflow Values")

#static plot
p_subset_altered

#interactive plot
#ggplotly(p_subset_altered)
```

# Water Year 2019

-   Read in the WY 2019 data from the External-MEF_DATA Box folder
    -   Raw data for the S2 Bog site is found at the following file
        path:
        `External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2`
-   Clean and manipulate the WY 2019 data into a tidy format
-   Change values of 0 to missing NA values during the following time
    periods:
    -   *April 1-2, Line and time is off*

    -   *April 8-17. Line doesn't fluctuate though there should've been
        rain or if not rain, it should've fallen from how high the flow
        is*

    -   *Proven in that the adjustment for 4/15 flow is great and S6
        Weir chart response. Therefore, no record*

    -   *April 17-22 discounted because pen level is off with no way to
        account for it*

    -   February: *Clock errors, record compromised. Flow is below zero
        for month Clock restarted 2/21/2019. Measurements still below
        zero.*

    -   January: *Clock errors meant that time could not be used to
        record accurate measurements. Measurements tend to be around
        0.001 and chart line is there or below zero. Physical
        measurements indicate flow until at least 1/15/2019*

    -   *March 14, Pen reverses and goes off the chart*

    -   *March 18th, Pen placed back on chart, still reversed (fixed
        4/23)*

    -   *March 30-31 Pen stays in place even though there should be a
        response. Likely because it reached the bottom of the chart (pen
        was reversed)*

    -   September: *Chart did not run from 9/3-9/10*
-   NA values are derived from notes recorded in .txt files found at the
    following Box filepath:
    `External-MEF_DATA/Hydro/Streamflow/L0_daily/StripCharts/DailyBreakpoint/2019/digitized/S2-19/`

```{r wy2019}
#create file path to call the data from Box 
## Mia's file path 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#read in the 2019 data 
wy2019 <- read_excel(path = here(filepath, "WY2019.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2019 data
wy2019_clean <- wy2019 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#assign February NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-01', '2019-04-02'))] <- NA

#assign March NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date == as.Date('2019-03-14')] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date == as.Date('2019-03-18')] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-03-30', '2019-03-31'))] <- NA

#assign April NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-01', '2019-04-02'))] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-08', '2019-04-17'))] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-17', '2019-04-22'))] <- NA

#assign September NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-09-03', '2019-09-10'))] <- NA
```

-   Test the data

```{r wy2019_test}
#test the 2019 data 
if(nrow(wy2019) != nrow(wy2019_clean)) stop("Check clean dataframe dimensions")
```

-   Plot the WY 2019 data

```{r wy2019_plot}
#sample size 
n <- nrow(wy2019_clean)

#plot 
p_2019 <- ggplot(data = wy2019_clean) + 
  geom_line(aes(x = datetime, 
                y = stream_height_ft, 
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3)))) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Stream Height (Ft)",
       title = "WY 2019 S2 Bog Stream Height") + 
  theme(plot.title = element_text(hjust = 0.5))

#static plot
#p_2019

#interactive plot
ggplotly(p_2019, tooltip = "text") %>% 
  layout(
    title = list(text = paste0("WY 2019 S2 Bog Weir Stream Height",
                                    '<br>',
                                    '<sup>',
                                     #subtitle
                                     paste0("n = ", n), 
                                    '</sup>')), 
         #title size 
         font = list(size = 14))
```

# Water Year 2020

-   Read in the WY 2020 data from the External-MEF_DATA Box folder
    -   Raw data for the S2 Bog site is found at the following file
        path:
        `External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2`
-   Clean and manipulate the WY 2020 data into a tidy format
-   Change values of 0 to missing NA values during the following time
    periods:
    -   *These gaps are not 0 flow (including the begin and end dates
        given), and should be replaced with NA unless Jake can digitize
        them: 7/28/2020 13:40:00 through 8/11/2020 00:00:00 (chart needs
        correction)*
-   NA values are derived from notes recorded in .txt files found at the
    following Box filepath:
    `External-MEF_DATA/Hydro/Streamflow/L0_daily/StripCharts/DailyBreakpoint/2020/digitized/S2-20/wy2020_s2_gaps.txt`

```{r wy2020}
#create file path to call the data from Box 
## Mia's file path 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#read in the 2020 data 
wy2020 <- read_excel(path = here(filepath, "WY2020.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2020 data
wy2020_clean <- wy2020 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#create date range to fill in with NA values 
start_date <- as.POSIXct("7/28/2020 13:40:00", 
                         format = '%m/%d/%Y %H:%M:%S',
                         tz = "GMT")
end_date <- as.POSIXct("8/11/2020 00:00:00", 
                         format = '%m/%d/%Y %H:%M:%S',
                         tz = "GMT")

#create a function to specify custom time intervals 
`%between%` <- function(x, interval) x >= interval[1] & x <= interval[2]

#assign February NA values 
wy2020_clean$stream_height_ft[wy2020_clean$datetime %between% c(start_date, end_date)] <- NA
```

-   Test the data

```{r wy2020_test}
#test the 2020 data 
if(nrow(wy2020) != nrow(wy2020_clean)) stop("Check clean dataframe dimensions")
```

-   Plot the WY 2020 data

```{r wy2020_plot}
#sample size
n <- nrow(wy2020_clean)

#plot 
p_2020 <- ggplot(data = wy2020_clean) + 
  geom_line(aes(x = datetime, 
                y = stream_height_ft,
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3)))) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Stream Height (Ft)",
       title = "WY 2020 S2 Bog Stream Height") + 
  theme(plot.title = element_text(hjust = 0.5))

#static plot
#p_2020

#interactive plot
ggplotly(p_2020, tooltip = "text") %>% 
  layout(
    title = list(text = paste0("WY 2020 S2 Bog Weir Stream Height",
                                    '<br>',
                                    '<sup>',
                                     #subtitle
                                     paste0("n = ", n), 
                                    '</sup>')), 
         #title size 
         font = list(size = 14))
```

# Water Year 2021

-   Read in the WY 2021 data from the External-MEF_DATA Box folder
    -   Raw data for the S2 Bog site is found at the following file
        path:
        `External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2`
-   Clean and manipulate the WY 2021 data into a tidy format
-   Change values of 0 to missing NA values during the following time
    periods:
    -   *These gaps are not 0 flow (including the begin and end dates
        given), and should be replaced with NA unless Jake can digitize
        them:*
        -   *2021-11-11 23:59:00 through 2021-12-29 00:00:00 (clock
            running fast on chart then clock died, well froze.
            Datalogger will have most of the data and is correct, except
            shaft encoder float froze at .14, see scanned chart for more
            remarks)*

        -   *and also (beginning values off, should be .18 float frozen
            for several hours) 2021-03-04 00:00:00 through 2021-03-09
            00:00:00 (need to correct for start of flow. See datalogger
            data)*
-   NA values are derived from notes recorded in .txt files found at the
    following Box filepath:
    `External-MEF_DATA/Hydro/Streamflow/L0_daily/StripCharts/DailyBreakpoint/2021/digitized/S2-21/wy2021_s2_gaps.txt`

```{r wy2021}
#create file path to call the data from Box 
## Mia's file path 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#read in the 2021 data 
wy2021 <- read_excel(path = here(filepath, "WY2021.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2021 data
wy2021_clean <- wy2021 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#identify date intervals that need to be filled w/NAs
date1 <- as.POSIXct("2021-03-04 00:00:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")
date2 <- as.POSIXct("2021-03-09 00:00:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")
date3 <- as.POSIXct("2021-11-11 23:59:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")
date4 <- as.POSIXct("2021-12-29 00:00:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")


#create a function to specify custom time intervals 
`%between%` <- function(x, interval) x >= interval[1] & x <= interval[2]

#assign March NA values
wy2021_clean$stream_height_ft[wy2021_clean$datetime %between% c(date1, date2)] <- NA

#assign Nov-Dec NA values 
wy2021_clean$stream_height_ft[wy2021_clean$datetime %between% c(date3, date4)] <- NA
```

-   Test the data

```{r wy2021_test}
#test the 2021 data 
if(nrow(wy2021) != nrow(wy2021_clean)) stop("Check clean dataframe dimensions")
```

-   Plot the WY 2021 data

```{r wy2021_plot}
#sample size
n <- nrow(wy2021_clean)

#plot 
p_2021 <- ggplot(data = wy2021_clean) + 
  geom_line(aes(x = datetime, 
                y = stream_height_ft,
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3)))) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Stream Height (Ft)",
       title = "WY 2021 S2 Bog Stream Height") + 
  theme(plot.title = element_text(hjust = 0.5))

#static plot
#p_2021

#interactive plot
ggplotly(p_2021, tooltip = "text") %>% 
  layout(
    title = list(text = paste0("WY 2021 S2 Bog Weir Stream Height",
                                    '<br>',
                                    '<sup>',
                                     #subtitle
                                     paste0("n = ", n), 
                                    '</sup>')), 
         #title size 
         font = list(size = 14))
```

# All Water Years (2017 - 2021)

-   Combine each individual water year

```{r all_streamflow}
#combine 2017 - 2021 data
all_streamflow <- rbind(wy2017_clean, 
                        wy2018_altered, 
                        wy2019_clean,
                        wy2020_clean,
                        wy2021_clean) %>% 
  #remove duplicate observations 
  unique()
```

-   Plot

```{r all_streamflow_plot}
#sample size 
n <- nrow(all_streamflow)

#plot 
p_all <- ggplot(data = all_streamflow) + 
  geom_line(aes(x = datetime, 
                y = stream_height_ft, 
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3)))) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Stream Height (Ft)",
       title = "S2 Bog Stream Height (WY 2017 - 2021)") + 
  theme(plot.title = element_text(hjust = 0.5))

#static plot
#p_all

#interactive plot
ggplotly(p_all, tooltip = "text") %>% 
  layout(
    title = list(text = paste0("S2 Bog Weir Stream Height (WY 2017 - 2021)",
                                    '<br>',
                                    '<sup>',
                                     #subtitle
                                     paste0("n = ", n), 
                                    '</sup>')), 
         #title size 
         font = list(size = 14))
```

Run this code chunk to knit to HTML without using the knit button in
RStudio

-   Save as `index.html` to render GitHub Page

```{r knit_to_html, eval=FALSE, include=FALSE}
#set file names and file path
input_file_name <- "1_water_years.Rmd"
output_file_name <- "index.html"
filepath <- "/Users/miaforsline/Desktop/water_years"

#knit to HTML
rmarkdown::render(input = input_file_name, 
                  output_format = "html_document", 
                  output_file = output_file_name, 
                  output_dir = filepath)
```

# Code Accessibility

Note that this file is being knit as
[`index.html`](https://github.com/miaforsline/water_years/blob/main/index.html)
into the [`water_years`
repository](https://github.com/miaforsline/water_years) in order to
update the [GitHub pages
website](https://miaforsline.github.io/water_years/), where the plots
can be viewed online.
