---
title: 'Marcell Experimental Forest: S2 Streamflow Manual Checkpoints'
author: "Mia Forsline"
date: '`r format(Sys.time(), "%B %d, %Y")`'
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, output_file="/Users/miaforsline/Desktop/manual_checks/index.html") })
output:
  html_document:
    theme: flatly
    code_folding: hide
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 72
---

# Objectives

The goal of this step in the workflow is to compare S2 stripchart stream
height values to the stream height values recorded at manual checkpoints
during the time period of interest. Manual checkpoints can be used to
quality check stream height values being collected by the stripchart.

# Set Up

-   Install/load necessary R packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE, 
                      warning = FALSE)

#install packages if necessary, then load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(
  here, #for file paths 
  janitor, 
  plotly, #for creating interactive plots 
  readxl, 
  tidyverse,
  zoo #for interpolating missing data points 
  )
```

-   Recreate intermediate data sets for use in this .Rmd file:
    -   Individual water years' worth of stripchart data

    -   All water years' worth of stripchart data

```{r intermediate_data}
#2017 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#read in the 2017 data 
wy2017 <- read_excel(path = here(filepath, "WY2017.S2_Breakpoint.xlsx"), 
                     skip =3) #skip the top 3 rows of the Excel Sheet 

#clean the 2017 data
wy2017_clean <- wy2017 %>% 
  #change column names
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  #remove unnecessary columns 
  select(c("datetime", "stream_height_ft")) %>% 
  #format column names into lower_snake_case
  clean_names() %>% 
  #create a date column 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#2018

filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#create a function to specify custom time intervals 
`%between%` <- function(x, interval) x >= interval[1] & x <= interval[2]

#read in the 2018 data 
wy2018 <- read_excel(path = here(filepath, "WY2018.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2018 data
wy2018_clean <- wy2018 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#assign October NA values 
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-10-09', '2018-10-10'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-10-11', '2018-10-15'))] <- NA

#assign November NA values 
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-11-10', '2018-11-14'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-11-25', '2018-12-04'))] <- NA

#assign December NA values 
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-12-01', '2018-12-03'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-12-08', '2018-12-10'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-12-11', '2018-12-31'))] <- NA

wy2018_altered <- wy2018_clean
wy2018_altered <- subset(wy2018_altered, !(wy2018_altered$date %between% as.Date(c('2018-10-31', '2018-11-01')) & wy2018_altered$stream_height_ft < 0.10))

filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#2019 
#read in the 2019 data 
wy2019 <- read_excel(path = here(filepath, "WY2019.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2019 data
wy2019_clean <- wy2019 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#assign February NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-01', '2019-04-02'))] <- NA

#assign March NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date == as.Date('2019-03-14')] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date == as.Date('2019-03-18')] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-03-30', '2019-03-31'))] <- NA

#assign April NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-01', '2019-04-02'))] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-08', '2019-04-17'))] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-17', '2019-04-22'))] <- NA

#assign September NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-09-03', '2019-09-10'))] <- NA

#2020 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#read in the 2020 data 
wy2020 <- read_excel(path = here(filepath, "WY2020.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2020 data
wy2020_clean <- wy2020 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#create date range to fill in with NA values 
start_date <- as.POSIXct("7/28/2020 13:40:00", 
                         format = '%m/%d/%Y %H:%M:%S',
                         tz = "GMT")
end_date <- as.POSIXct("8/11/2020 00:00:00", 
                         format = '%m/%d/%Y %H:%M:%S',
                         tz = "GMT")

#create a function to specify custom time intervals 
`%between%` <- function(x, interval) x >= interval[1] & x <= interval[2]

#assign February NA values 
wy2020_clean$stream_height_ft[wy2020_clean$datetime %between% c(start_date, end_date)] <- NA

#2021 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#read in the 2021 data 
wy2021 <- read_excel(path = here(filepath, "WY2021.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2021 data
wy2021_clean <- wy2021 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#identify date intervals that need to be filled w/NAs
date1 <- as.POSIXct("2021-03-04 00:00:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")
date2 <- as.POSIXct("2021-03-09 00:00:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")
date3 <- as.POSIXct("2021-11-11 23:59:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")
date4 <- as.POSIXct("2021-12-29 00:00:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")


#create a function to specify custom time intervals 
`%between%` <- function(x, interval) x >= interval[1] & x <= interval[2]

#assign March NA values
wy2021_clean$stream_height_ft[wy2021_clean$datetime %between% c(date1, date2)] <- NA

#assign Nov-Dec NA values 
wy2021_clean$stream_height_ft[wy2021_clean$datetime %between% c(date3, date4)] <- NA

#all water years 
all_streamflow <- rbind(wy2017_clean, 
                        wy2018_altered, 
                        wy2019_clean,
                        wy2020_clean,
                        wy2021_clean)
```

-   Set date range

```{r min_max_year}
min_year = 2017
max_year = 2021
```

# Manual Streamflow Checkpoints (2017 - 2021)

-   Read in the manual streamflow checkpoints from the External-MEF_DATA
    Box folder
    -   The raw data is found at the following file path:
        `External-MEF_DATA/Hydro/Streamflow/L0_subdaily/ManualChecks`
-   Clean and manipulate the data into a tidy format
    -   Renaming columns to use lower_snake_case and avoid special
        characters via `clean_names()`
    -   Convert the `collected` column from a character into a POSIXct
        date format
-   Subset the data to include only the data of interest (S2 weir data)
    and the time period of interest (2017 - 2021)

```{r manual_checks}
#create file path to call the data from Box 
## Mia's file path 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/ManualChecks"

#read in the manual checks data 
mc <- read_csv(here(filepath, "2017-2021_S2Stage.csv"))

#clean the data
mc_clean <- mc %>% 
  clean_names() %>% 
  #remove S2 lagging pool data and keep only the S2 weir data
  subset(name == "S2 WEIR") %>%
  mutate(collected = as.POSIXct(collected, format = '%m/%d/%Y %H:%M', 
                                tz = "GMT"),
         year = format(as.POSIXct(collected, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         date = format(as.POSIXct(collected, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         diff = point_gage - stripchart_stage
         ) %>% 
  subset(year >= min_year & year <= max_year) 

#create subsets of the data where NA values are removed from a col of interest
#point check data 
pc_clean <- mc_clean %>% 
  #remove rows with NA values in the point_gage col 
  subset(!is.na(point_gage))

#stripchart check data 
sc_clean <- mc_clean %>% 
  #remove rows with NA values in the stripchart_stage col 
  subset(!is.na(stripchart_stage))

#format data for plotting 
mc_plot <- mc_clean %>% 
  select(-logger_stage,
         -year,
         -date,
         -site,
         -name) %>% 
  pivot_longer(cols = c("point_gage", "stripchart_stage"),
               names_to = "type", 
               values_to = "stream_height_ft")

```

-   Plot the manual checks data

```{r plot}
#sample size
n <- nrow(mc_plot)

ggplot(data = mc_plot) + 
  geom_point(aes(x = collected, 
                 y = stream_height_ft,
                 color = type),
             alpha = 0.5) +  
  theme_classic() + 
  labs(y = "Stream Height (ft)", 
       x = "Time", 
       title = paste0("S2 Manual Checkpoints VS \n Stripchart Values Over Time (", min_year, "-", max_year, ")"),
       color = "Checkpoint Type",
       caption = paste0("n = ", n)
       ) + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_discrete(labels = c("Point Gage", "Stripchart Stage"))
```

-   Plot the differences over time, where the difference is calculated
    as `point_gage` - `stripchart_stage`

```{r plot_diff}
#sample size
n <- nrow(mc_plot)

ggplot(data = mc_plot) + 
  geom_point(aes(x = collected, y = diff)) + 
  geom_smooth(aes(x = collected, y = diff), method = "lm") +  
  theme_classic() + 
  labs(y = "Stream Height (ft)", 
       x = "Point Gage - Stripchart Stage", 
       title = paste0("S2 Manual Checkpoints VS \n Stripchart Differences Over Time (", min_year, "-", max_year, ")"), 
       caption = paste0("n = ", n)
       ) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) + 
  scale_color_discrete(labels = c("Point Gage", "Stripchart Stage"))
```

# Streamflow + Manual Checks (2017 - 2021)

-   Plot stripchart data (as lines) and manual checkpoints (as points)

For interactive plots,

-   stripchart stream height (ft) is plotted as a **black line**, and

-   manual checkpoints of stream height (ft) are plotted as [red
    dots]{style="color:red"}.

```{r wy2017-2021}
#plot
p_all <- ggplot() +
  #stripchart streamflow data
  geom_line(data = all_streamflow,
            aes(x = datetime, 
                y = stream_height_ft,
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3))),
            size = 0.25) + 
  #manual checkpoints 
  geom_point(data = mc_clean, 
             aes(x = collected, 
                 y = point_gage,
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", collected, "\n",
                              "Stream Height Checkpoint (ft): ", round(stripchart_stage, digits = 3))), 
             color = "red",
             size = 0.5) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Stream Height (ft)",
       title = paste0("S2 Bog Stream Height (", min_year, "-", max_year, ")"
       ),
       subtitle = "Stripchart data are plotted as black lines. Manual checks are plotted as red dots.") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

#static plot
#p_all

#interactive plot 
ggplotly(p_all, tooltip = "text")
```

## 1:1 Stripchart vs Manual Checkpoints Comparison (2017 - 2021)

Next, we are interested in a 1:1 comparison of stripchart data vs manual
checkpoints data at the exact same timestamp. Since the stripchart data
and manual checkpoints do not align perfectly, we will interpolate the
stripchart stream flow values using the `na.approx()` function from the
`zoo` package to estimate stripchart values at the time of the manual
checks.

## Stripchart vs Stripchart Stage

-   Clean the data

-   Note that the manual checks data only range from 2017-04-04 to
    2019-12-31, which is a smaller time range than the stripchart data

```{r interpolate_sc}
#clean the stripchart_stage data (using the point_gage col)
sc_sub <- sc_clean %>% 
  #extract the data (without the timestamp)
  mutate(date = format(as.POSIXct(collected, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT"),
         year = as.numeric(year)
         ) %>% 
  #rename column
  rename(datetime = collected) %>% 
  #remove unnecessary columns 
  select(-site, -lab_id, -name, -point_gage, -logger_stage, -diff) 

#identify date ranges of interest: 
##aka the range of the manual checkpoint data
max_date <- max(sc_sub$date)
min_date <- min(sc_sub$date)

#subset stripchart data to fit within the time range of the manual checkpoints
streamflow_sub <- all_streamflow %>% 
  mutate(
    year = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                             tz = "GMT"),
                  format = '%Y'), 
    year = as.numeric(year)
    ) %>% 
  subset(date <= max_date & date >= min_date) 
```


```{r na.approx()}
test <- full_join(x = sc_sub, 
                y = streamflow_sub, 
                by = c("datetime", "date", "year")) %>% 
  arrange(datetime)

x <- test[14:16, ]

y <- x %>% 
  mutate(approx = na.approx(stream_height_ft, 
                            na.rm =FALSE, 
                            method = "linear", 
                            maxgap = 6)
         )
```


-   Join the manual checkpoints (`sc_sub`) dataset with the
    continuous stripchart dataset (`streamflow_sub`)

-   Optional: Plot the data to visually examine if the approximated values
    generated by na.approx() look correct

```{r join_sc}
#join the stripchart data and manual checks data 
fj_sc <- full_join(x = sc_sub, 
                y = streamflow_sub, 
                by = c("datetime", "date", "year"))  %>%
  #interpolate to fill in missing NA values
  mutate(approx = na.approx(object = stream_height_ft,
                            #x = datetime,
                            #xout = datetime,
                            method = "linear",
                            maxgap = 6),
         #calculate the difference between approximated values and real values
         diff = approx - stream_height_ft) %>%
  #rearrange dataframe into a long format
  pivot_longer(cols = c("stripchart_stage", "stream_height_ft"),
               names_to = "types",
               values_to = "stream_height_ft")

#test the full join did not drop any observations 
if(nrow(fj_sc) != (nrow(sc_sub) + nrow(streamflow_sub))) stop("Check lj dataframe dimensions")

##visually examine the approximated values 
# ggplot(data = fj_sc) +
#   geom_line(aes(x = datetime, y = approx))
# 
# ggplot(data = fj_sc) +
#   geom_point(aes(x = datetime,
#                 y = stream_height_ft,
#                 color = types))
# 
# ggplot(data = fj_sc) +
#   geom_point(aes(x = datetime,
#                 y = stream_height_ft,
#                 color = types)) +
#   geom_point(aes(x = datetime,
#                 y = approx),
#              alpha = 0.05)


fj_sc_wide <- fj_sc %>%
  #remove extraneous columns 
  select(-stream_height_ft, -diff) %>% 
  #return to wide format to create scatterplot
  pivot_wider(
    names_from = "types",
    values_from = "approx"
  ) %>% 
  #unlist the columns created by pivot_wider() 
  unnest()


# #visually examine all data 
# ggplot(data = fj_sc) +
#   geom_point(aes(x = manual_check,
#                  y = stripchart))
```

-   Subset the joined dataframe to include only the timestamps of
    interest (AKA the timestamps from the original stripchart_stage
    dataset)

-   Plot

```{r 1_to_1_sc}
#left join the joined data and stripchart_stage data to keep only the timestamps of interest 
lj_sc <- left_join(x = sc_sub, 
                y = fj_sc,
                by = c("datetime", "year", "stripchart_stage", "date")) %>% 
  #ensure the correct column types
  mutate(stripchart_stage = as.numeric(stripchart_stage),
         stream_height_ft = as.numeric(stream_height_ft),
         diff = stripchart_stage - stream_height_ft)

#test if the subsetted data has the same number of observations as the manual checks dataframe
#if(nrow(lj) != (nrow(pc_sub))) stop("Check lj dataframe dimensions")

#test if the approximated data differs greatly from the stripchart/manual checks data 
#if(abs(lj$diff) > 0.01 ) stop("Check differences")
```

```{r 1_to_1_plot_sc}
#plot 
p_1to1 <- ggplot() + 
  geom_point(data = lj_sc, 
             aes(x = stripchart_stage, #approximated values from manual checkpoints
                 y = stream_height_ft, #continuous stripchart values 
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("Stripchart Stage (ft): ", stripchart_stage, "\n",
                              "Stripchart (ft): ", stream_height_ft)),
             alpha = 0.4) + 
  theme_classic() +
  labs(x = "Stripchart Stage",
       y = "Stripchart",
       title = paste0("S2 Stripchart vs Interpolated Stripchart Values (", min_year, "-", max_year, ")"),
       subtitle = "The line y = x is plotted for reference.") +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) + 
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.4) +
  ylim(0, 0.4)

#static plot
#p_1to1

#interactive plot 
ggplotly(p_1to1, tooltip = "text")
```

## Stripchart vs Point Gage Manual Checkpoints

```{r pc}
#clean the point gage data 
pc_sub <- pc_clean %>% 
  #extract the data (without the timestamp)
  mutate(date = format(as.POSIXct(collected, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT"),
         year = as.numeric(year)
         ) %>% 
  #rename column
  rename(datetime = collected) %>% 
  #remove unnecessary columns 
  select(-site, 
         -lab_id, 
         -name, 
         -stripchart_stage, 
         -logger_stage,
         -diff)
  

#identify date ranges of interest: 2017-04-04 to 2019-12-31
##aka the range of the manual checkpoint data
max_date <- max(pc_sub$date)
min_date <- min(pc_sub$date)
nrows <- nrow(pc_sub)

# #subset stripchart data to fit within the time range of the manual checkpoints
# streamflow_sub <- all_streamflow %>% 
#   mutate(
#     year = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
#                              tz = "GMT"),
#                   format = '%Y'), 
#     year = as.numeric(year)
#     ) %>% 
#   subset(date <= max_date & date >= min_date) 
```

-   Join the manual checks dataset with the stripchart dataset

-   Plot the data to visually examine if the approximated values
    generated by na.approx() look correct

```{r join_pc}
#join the stripchart data and manual checks data 
fj_pc <- full_join(x = pc_sub, 
                y = streamflow_sub, 
                by = c("datetime", "date", "year")) %>% 
  #rename columns 
  rename(manual_check = point_gage,
         stripchart = stream_height_ft) %>% 
  #interpolate to fill in missing NA values
  mutate(approx = na.approx(object = stripchart,
                            x = datetime,
                            method = "linear",
                            maxgap = 6))

# #rearrange dataframe into a long format
#   pivot_longer(cols = c("manual_check", "stripchart"),
#                names_to = "types",
#                values_to = "stream_height_ft") %>%

# #visually examine the approximated values
# ggplot(data = fj_pc) +
#   geom_line(aes(x = datetime, y = approx))
# 
# ggplot(data = fj_pc) +
#   geom_point(aes(x = datetime,
#                 y = stream_height_ft,
#                 color = types))
# 
# ggplot(data = fj_pc) +
#   geom_point(aes(x = datetime,
#                 y = stream_height_ft,
#                 color = types)) +
#   geom_point(aes(x = datetime,
#                 y = approx),
#              alpha = 0.05)


fj_pc <- fj_pc %>%
  #remove extraneous columns 
  select(-stripchart) 
# %>% 
#   #return to wide format to create scatterplot
#   pivot_wider(
#     names_from = "types",
#     values_from = "approx"
#   ) %>% 
#   #unlist the columns created by pivot_wider() 
#   unnest


# #visually examine all data
# ggplot(data = fj_pc) +
#   geom_point(aes(x = manual_check,
#                  y = stripchart))
```

-   Subset the joined dataframe to include only the timestamps of
    interest (AKA the timestamps from the original manual checks
    dataset)

-   Plot

```{r 1_to_1_pc}
#left join the joined data and manual checks data to keep only the timestamps of interest 
lj_pc <- left_join(x = pc_sub, 
                y = fj_pc,
                by = c("datetime", "year", "date", "point_gage" = "manual_check")) %>% 
  #ensure the correct column types
  mutate(point_gage = as.numeric(point_gage),
         approx = as.numeric(approx),
         diff = point_gage - approx)

#test if the subsetted data has the same number of observations as the manual checks dataframe
#if(nrow(lj_pc) != (nrow(pc_sub))) stop("Check lj dataframe dimensions")

#test if the approximated data differs greatly from the stripchart/manual checks data 
#if(abs(lj_pc$diff) > 0.01 ) stop("Check differences")
```

```{r 1_to_1_plot_pc}
#plot 
p_1to1 <- ggplot() + 
  geom_point(data = lj_pc, 
             aes(x = point_gage, 
                 y = approx, 
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("Manual Checkpoint (ft): ", point_gage, "\n",
                              "Stripchart (ft): ", approx)),
             alpha = 0.5) + 
  theme_classic() +
  labs(x = "Manual Checkpoints",
       y = "Stripchart Data",
       title = paste0("S2 Manual Checks vs Interpolated Stripchart Values (", min_year, "-", max_year, ")"),
       subtitle = "The line y = x is plotted for reference.") +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) + 
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.4) +
  ylim(0, 0.4)

#static plot
#p_1to1

#interactive plot 
ggplotly(p_1to1, tooltip = "text")
```

Run this code chunk to knit to HTML without using the knit button in
RStudio

-   Save as `index.html` to render GitHub Page

```{r knit_to_html, eval=FALSE, include=FALSE}
#set file names and file path
input_file_name <- "2_manual_checks.Rmd"
output_file_name <- "index.html"
filepath <- "/Users/miaforsline/Desktop/manual_checks"

#knit to HTML
rmarkdown::render(input = input_file_name, 
                  output_format = "html_document", 
                  output_file = output_file_name, 
                  output_dir = filepath)
```

# Code Accessibility

Note that this file is being knit as
[`index.html`](https://github.com/miaforsline/manual_checks/blob/main/index.html)
into the [`manual_checks`
repository](https://github.com/miaforsline/manual_checks) in order to
update the [GitHub pages
website](https://miaforsline.github.io/manual_checks/), where the plots
can be viewed online.
