---
title: 'Marcell Experimental Forest: S2 Climate and Bogwell Data'
author: "Mia Forsline"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    theme: flatly
    code_folding: hide
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 72
---

# Set Up

Install/load necessary R packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE, 
                      warning = FALSE)

#install packages if necessary, then load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(
  here, #for file paths 
  janitor, 
  plotly, #for creating interactive plots 
  readxl, 
  tidyverse,
  zoo #for interpolating missing data points 
  )
```

Set date range

```{r min_max_year}
min_year = 2017
max_year = 2021
```

Create sub-directories if necessary 

```{r directories}
output_dir <- file.path(here("intermediate_data"))

if (!dir.exists(output_dir)){
dir.create(output_dir)
} else {
    print("Directory already exists!")
}

output_dir <- file.path(here("figures"))

if (!dir.exists(output_dir)){
dir.create(output_dir)
} else {
    print("Directory already exists!")
}
```

# Climate and Bogwell data

# Precipitation 

## Precipitation Every 15min

-   Read in the data from the following Box file path:
    `External-MEF_DATA/EDI/precipitation_15min/edi.849.2/data_objects/`

-   Clean the data into a tidy format

    -   Subset precipitation data to 2017 - present

-   Plot

```{r precip_15}
#set file path to read in data from Box 
filepath = "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/EDI/precipitation_15min/edi.849.2/data_objects/"

#read in the raw data 
precip_15 <- read_csv(here(filepath, "MEF_South_precip_15min.csv"))

#clean the data 
precip_15_clean <- precip_15 %>% 
  clean_names() %>% 
  mutate(year = format(as.POSIXct(timestamp, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year)) %>% 
  subset(year >= min_year) %>% 
  rename(precip_in = south_pcp) %>% 
  mutate(precip_10 = (precip_in / 10)) %>% 
  pivot_longer(
    cols = c("precip_in", "precip_10"),
    names_to = "precip_type",
    values_to = "precip_in"
  ) %>% 
  rename(datetime = timestamp)

#save clean data CSV to use in future RMD files
write.csv(x = precip_15_clean, 
          file = file.path(here("intermediate_data", "precip_15_clean.csv")),
          row.names = FALSE)

#plot 
p_15 <- ggplot(data = precip_15_clean) + 
  geom_line(aes(x = datetime, y = precip_in, color = precip_type)) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Precipitation (cm)", 
       title = paste0("Precipitation Sum Every 15 Min (", min_year, "-", max_year, ")"))

#static plot
p_15

#interactive plot 
#ggplotly(p_15)
```

## Precipitation Every 6 Hours

```{r precip_6h}
#read in 15min precip data 
precip_6h <- precip_15 

#clean and sum precip by 6 hours 
precip_6h_clean <- precip_6h %>% 
  clean_names() %>% 
  mutate(timestamp = lubridate::floor_date(timestamp, unit = "6 hours")) %>%
    group_by(timestamp) %>%
    dplyr::summarize(south_pcp = sum(south_pcp)) %>%
    ungroup() %>% 
  rename(datetime = timestamp,
         precip_in = south_pcp) %>% 
  mutate(
    year = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year),
     date = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
    precip_10_in = precip_in / 10
  ) %>% 
  subset(year >= min_year & year <= max_year)

#save clean data CSV to use in future RMD files
write.csv(x = precip_6h_clean, 
          file = file.path(here("intermediate_data", "precip_6h_clean.csv")),
          row.names = FALSE)

#plot
p_precip_6h <- ggplot(data = precip_6h_clean) + 
  geom_col(aes(x = datetime, y = precip_in)) + 
  theme_classic() + 
  labs(x = "Time",
       y = "Cumulative Precipitation (in)", 
       title = paste0("Precipitation Sum Every 6 Hours (", min_year, "-", max_year, ")")) + 
  theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))

#static plot
p_precip_6h

#interactive plot
#ggplotly(p_precip_6h)

```



## Precipitation Every 12 Hours

```{r precip_12h}
#read in 15min precip data 
precip_12h <- precip_15 

#clean and sum precip by 6 hours 
precip_12h_clean <- precip_12h %>% 
  clean_names() %>% 
  mutate(timestamp = lubridate::floor_date(timestamp, unit = "12 hours")) %>%
    group_by(timestamp) %>%
    dplyr::summarize(south_pcp = sum(south_pcp)) %>%
    ungroup() %>% 
  rename(datetime = timestamp,
         precip_in = south_pcp) %>% 
  mutate(
    year = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year),
    date = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
    precip_10_in = precip_in / 10
  ) %>% 
  subset(year >= min_year & year <= max_year)

#save clean data CSV to use in future RMD files
write.csv(x = precip_12h_clean, 
          file = file.path(here("intermediate_data", "precip_12h_clean.csv")),
          row.names = FALSE)

#plot
p_precip_12h <- ggplot(data = precip_12h_clean) + 
  geom_col(aes(x = datetime, y = precip_in)) + 
  theme_classic() + 
  labs(x = "Time",
       y = "Cumulative Precipitation (in)", 
       title = paste0("Precipitation Sum Every 12 Hours (", min_year, "-", max_year, ")")) + 
  theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))

#static plot 
p_precip_12h

#interactive plot
#ggplotly(p_precip_12h)
```


## Precipitation Daily

-   Read in the data from the following Box file path:
    `External-MEF_DATA/EDI/precipitation_daily/edi.563.3/data_objects/`

-   Clean the data into a tidy format

    -   Subset precipitation data to 2017 - present

-   Plot

```{r precip_daily}
#set file path to read in data from Box 
filepath = "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/EDI/precipitation_daily/edi.563.3/data_objects/"

#read in the raw data 
precip_daily <- read_csv(here(filepath, "MEF_precipitation_daily.csv"))

#clean the data 
precip_daily_clean <- precip_daily %>% 
  clean_names() %>% 
  select(date, south_pcp, south_flag) %>% 
  mutate(year = format(as.POSIXct(date, format = '%Y-%m-%d', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year)) %>% 
  subset(year >= min_year) %>% 
  rename(precip_in = south_pcp) %>% 
  mutate(precip_10 = (precip_in / 10)) %>% 
  pivot_longer(
    cols = c("precip_in", "precip_10"),
    names_to = "precip_type",
    values_to = "precip_in"
  )

#save clean data CSV to use in future RMD files
write.csv(x = precip_daily_clean, 
          file = file.path(here("intermediate_data", "precip_daily_clean.csv")),
          row.names = FALSE)

#plot 
p_daily <- ggplot(data = precip_daily_clean) + 
  geom_line(aes(x = date, y = precip_in, color = precip_type)) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "South Precipitation (in)", 
       title = paste0("Daily Cumulative Precipitation (", min_year, "-", max_year, ")")
       )

#static plot
p_daily

#ggplotly(p_daily)

#Nina's code for calculating daily precip sums 
# precipDaily <- precip15 %>%
#   mutate(DATE = lubridate::floor_date(TIMESTAMP, unit = "day")) %>%
#   group_by(DATE) %>%
#   dplyr::summarize(South = sum(South)) %>%
#   ungroup()
```

# Air Temperature Every 30 Min

-   Read in the data from the following Box file path:
    `External-MEF_DATA/EDI/airT_RH_BP_30min/edi.990.1/data_objects/`

-   Clean the data into a tidy format

    -   Subset precipitation data to 2017 - present

-   Plot

```{r airtemp_30}
#set file path to read in data from Box 
filepath = "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/EDI/airT_RH_BP_30min/edi.990.1/data_objects/"

#read in the raw data 
air_temp_30 <- read_csv(here(filepath, "MEF_S2Bog_Met_30min.csv"))

#clean the data 
air_temp_30_clean <- air_temp_30 %>% 
  clean_names() %>% 
  rename(airtempc = air_t,
         rel_humidity = rh,
         datetime = timestamp) %>% 
  mutate(year = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year),
         airtempc_10 = (airtempc / 10),
         airtempc_100 = (airtempc / 100),
         #create a variable that indicates if the airtemp is above or below freezing
         freezing = 0
         ) %>% 
  pivot_longer(
    cols = c("airtempc", "airtempc_10", "airtempc_100"),
    names_to = "airtemp_type",
    values_to = "airtempC"
  )

#assign a value of 1 to the variable if airtemp is above freezing 
air_temp_30_clean$freezing[air_temp_30_clean$airtempC > 0] <- 1

#save clean data CSV to use in future RMD files
write.csv(x = air_temp_30_clean, 
          file = file.path(here("intermediate_data", "air_temp_30_clean.csv")),
          row.names = FALSE)

#plot 
p_air_30 <- ggplot(data = air_temp_30_clean) + 
  geom_line(aes(x = datetime, y = airtempC, color = airtemp_type)) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Air Temperature (ºC)", 
       title = paste0("Air Temperature Every 30 Min (", min_year, "-", max_year, ")") 
       ) + 
  theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))

#static plot
p_air_30

#interactive plot
#ggplotly(p_air_30)
```

# Peatland Daily Water Table

-   Read in the data from the following Box file path:
    `External-MEF_DATA/EDI/peatland_water_table_daily/edi.562.2/data_objects/`

-   Clean the data into a tidy format

    -   Subset precipitation data to 2017 - present

-   Plot

```{r bogwell_daily}
#set file path to read in data from Box 
filepath = "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/EDI/peatland_water_table_daily/edi.562.2/data_objects/"

#read in the raw data 
bogwell <- read_csv(here(filepath, "MEF_daily_peatland_water_table.csv"))

#clean the data
bogwell_clean <- bogwell %>% 
  clean_names() %>% 
  subset(peatland == "S2") %>% 
  mutate(year = format(as.POSIXct(date, format = '%Y-%m-%d', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year)) %>% 
  subset(year >= min_year & year <= max_year)

#save clean data CSV to use in future RMD files
write.csv(x = bogwell_clean, 
          file = file.path(here("intermediate_data", "bogwell_clean.csv")),
          row.names = FALSE)

#plot
p_bogwell <- ggplot(data = bogwell_clean) + 
  geom_line(aes(x = date, y = wte)) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Water Table Elevation (m above sea level)",
       title = paste0("S2 Daily Peatland Water Table Elevation (", min_year, "-", max_year, ")") 
       )

#static plot 
p_bogwell

#interactive plot
#ggplotly(p_bogwell)
```

