---
title: 'Marcell Experimental Forest: S2 Climate Data'
author: "Mia Forsline"
date: '`r format(Sys.time(), "%B %d, %Y")`'
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, output_file="/Users/miaforsline/Desktop/climate/index.html") })
output:
  html_document:
    theme: flatly
    code_folding: hide
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 72
---

# Objectives

The goal of this step in the workflow is to import, clean, and visualize
relevant climate data (precipitation and air temperature) to later
compare to streamflow height values collected from the Marcell
Experimental Forest.

# Set Up

-   Install/load necessary R packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE, 
                      warning = FALSE)

#install packages if necessary, then load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(
  here, #for file paths 
  janitor, 
  kableExtra,
  plotly, #for creating interactive plots 
  readxl, 
  tidyverse,
  zoo #for interpolating missing data points 
  )
```

-   Set date range

```{r min_max_year}
min_year = 2017
max_year = 2021
```

# Precipitation

More information about precipitation data collected at the Marcell
Experimental Forest can be found on the [Environmental Data Initiative
Repository
Portal](https://portal.edirepository.org/nis/mapbrowse?scope=edi&identifier=563).

## Cumulative Precipitation Every 15min

-   Read in the data from the following Box file path:
    `External-MEF_DATA/EDI/precipitation_15min/edi.849.2/data_objects/`

-   Clean the data into a tidy format

    -   Subset precipitation data to 2017 - present

-   Plot

```{r precip_15}
#set file path to read in data from Box 
filepath = "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/EDI/precipitation_15min/edi.849.2/data_objects/"

#read in the raw data 
precip_15 <- read_csv(here(filepath, "MEF_South_precip_15min.csv"))

#clean the data 
precip_15_clean <- precip_15 %>% 
  clean_names() %>% 
  mutate(year = format(as.POSIXct(timestamp, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year)) %>% 
  subset(year >= min_year) %>% 
  rename(precip_in = south_pcp) %>% 
  mutate(precip_10 = (precip_in / 10)) %>% 
  pivot_longer(
    cols = c("precip_in", "precip_10"),
    names_to = "precip_type",
    values_to = "precip_in"
  ) %>% 
  rename(datetime = timestamp)

#plot 
p_15 <- ggplot(data = precip_15_clean) + 
  geom_line(aes(x = datetime, 
                y = precip_in, 
                color = precip_type),
            alpha = 0.5) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Precipitation (cm)", 
       title = paste0("Precipitation Sum Every 15 Min (", min_year, "-", max_year, ")"),
       color = "Precipitation Type") + 
  scale_color_discrete(labels = c("inches / 10", "inches"))

#static plot
p_15

#interactive plot 
#ggplotly(p_15)
```

## Cumulative Precipitation Every 6 Hours

```{r precip_6h}
#read in 15min precip data 
precip_6h <- precip_15 

#clean and sum precip by 6 hours 
precip_6h_clean <- precip_6h %>% 
  clean_names() %>% 
  mutate(timestamp = lubridate::floor_date(timestamp, unit = "6 hours")) %>%
    group_by(timestamp) %>%
    dplyr::summarize(south_pcp = sum(south_pcp)) %>%
    ungroup() %>% 
  rename(datetime = timestamp,
         precip_in = south_pcp) %>% 
  mutate(
    year = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year),
     date = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
    precip_10_in = precip_in / 10
  ) %>% 
  subset(year >= min_year & year <= max_year)

#plot
p_precip_6h <- ggplot(data = precip_6h_clean) + 
  geom_col(aes(x = datetime, y = precip_in)) + 
  theme_classic() + 
  labs(x = "Time",
       y = "Cumulative Precipitation (in)", 
       title = paste0("Precipitation Sum Every 6 Hours (", min_year, "-", max_year, ")")) + 
  theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))

#static plot
p_precip_6h

#interactive plot
#ggplotly(p_precip_6h)

```

## Cumulative Precipitation Every 12 Hours

```{r precip_12h}
#read in 15min precip data 
precip_12h <- precip_15 

#clean and sum precip by 6 hours 
precip_12h_clean <- precip_12h %>% 
  clean_names() %>% 
  mutate(timestamp = lubridate::floor_date(timestamp, unit = "12 hours")) %>%
    group_by(timestamp) %>%
    dplyr::summarize(south_pcp = sum(south_pcp)) %>%
    ungroup() %>% 
  rename(datetime = timestamp,
         precip_in = south_pcp) %>% 
  mutate(
    year = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year),
    date = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
    precip_10_in = precip_in / 10
  ) %>% 
  subset(year >= min_year & year <= max_year)

#plot
p_precip_12h <- ggplot(data = precip_12h_clean) + 
  geom_col(aes(x = datetime, y = precip_in)) + 
  theme_classic() + 
  labs(x = "Time",
       y = "Cumulative Precipitation (in)", 
       title = paste0("Precipitation Sum Every 12 Hours (", min_year, "-", max_year, ")")) + 
  theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))

#static plot 
p_precip_12h

#interactive plot
#ggplotly(p_precip_12h)
```

## Cumulative Precipitation Daily

-   Read in the data from the following Box file path:
    `External-MEF_DATA/EDI/precipitation_daily/edi.563.3/data_objects/`

-   Clean the data into a tidy format

    -   Subset precipitation data to 2017 - present

-   Plot

```{r precip_daily}
#set file path to read in data from Box 
filepath = "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/EDI/precipitation_daily/edi.563.3/data_objects/"

#read in the raw data 
precip_daily <- read_csv(here(filepath, "MEF_precipitation_daily.csv"))

#clean the data 
precip_daily_clean <- precip_daily %>% 
  clean_names() %>% 
  select(date, south_pcp, south_flag) %>% 
  mutate(year = format(as.POSIXct(date, format = '%Y-%m-%d', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year)) %>% 
  subset(year >= min_year) %>% 
  rename(precip_in = south_pcp) %>% 
  mutate(precip_10 = (precip_in / 10)) %>% 
  pivot_longer(
    cols = c("precip_in", "precip_10"),
    names_to = "precip_type",
    values_to = "precip_in"
  )

#plot 
p_daily <- ggplot(data = precip_daily_clean) + 
  geom_line(aes(x = date, 
                y = precip_in, 
                color = precip_type),
            alpha = 0.5) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "South Precipitation (in)", 
       title = paste0("Daily Cumulative Precipitation (", min_year, "-", max_year, ")"),
       color = "Precipitation Type"
       ) + 
  scale_color_discrete(labels = c("inches / 10", "inches"))

#static plot
p_daily

#ggplotly(p_daily)

#Nina's code for calculating daily precip sums 
# precipDaily <- precip15 %>%
#   mutate(DATE = lubridate::floor_date(TIMESTAMP, unit = "day")) %>%
#   group_by(DATE) %>%
#   dplyr::summarize(South = sum(South)) %>%
#   ungroup()
```

# Air Temperature Every 30 Min

-   Read in the data from the following Box file path:
    `External-MEF_DATA/EDI/airT_RH_BP_30min/edi.990.1/data_objects/`

-   Clean the data into a tidy format

    -   Subset precipitation data to 2017 - present

-   Plot

```{r airtemp_30}
#set file path to read in data from Box 
filepath = "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/EDI/airT_RH_BP_30min/edi.990.1/data_objects/"

#read in the raw data 
air_temp_30 <- read_csv(here(filepath, "MEF_S2Bog_Met_30min.csv"))

#clean the data 
air_temp_30_clean <- air_temp_30 %>% 
  clean_names() %>% 
  rename(airtempc = air_t,
         rel_humidity = rh,
         datetime = timestamp) %>% 
  mutate(year = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year),
         airtempc_10 = (airtempc / 10),
         airtempc_100 = (airtempc / 100),
         #create a variable that indicates if the airtemp is above or below freezing
         freezing = 0
         ) %>% 
  pivot_longer(
    cols = c("airtempc", "airtempc_10", "airtempc_100"),
    names_to = "airtemp_type",
    values_to = "airtempC"
  )

#assign a value of 1 to the variable if airtemp is above freezing 
air_temp_30_clean$freezing[air_temp_30_clean$airtempC > 0] <- 1

#plot 
p_air_30 <- ggplot(data = air_temp_30_clean) + 
  geom_line(aes(x = datetime, 
                y = airtempC, 
                color = airtemp_type)) + 
  theme_classic() + 
  labs(x = "Time", 
       y = "Air Temperature (ºC)", 
       title = paste0("Air Temperature Every 30 Min (", min_year, "-", max_year, ")"), 
       color = "Air Temperature Type"
       ) + 
  theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) + 
  scale_color_discrete(labels = c("ºC", 
                                  "ºC / 10", 
                                  "ºC / 100"))

#static plot
p_air_30

#interactive plot
#ggplotly(p_air_30)
```

Run this code chunk to knit to HTML without using the knit button in
RStudio

-   Save as `index.html` to render GitHub Page

```{r knit_to_html, eval=FALSE, include=FALSE}
#set file names and file path
input_file_name <- "3_climate.Rmd"
output_file_name <- "index.html"
filepath <- "/Users/miaforsline/Desktop/climate"

#knit to HTML
rmarkdown::render(input = input_file_name, 
                  output_format = "html_document", 
                  output_file = output_file_name, 
                  output_dir = filepath)
```

# Code Accessibility

Note that this file is being knit as
[`index.html`](https://github.com/miaforsline/climate/blob/main/index.html)
into the public [`climate` GitHub
repository](https://github.com/miaforsline/climate) in order to update
the [GitHub pages website](https://miaforsline.github.io/climate/),
where the plots can be viewed online.
