---
title: 'Marcell Experimental Forest: S2 Stream Height & Precipitation Comparison Analyses'
author: "Mia Forsline"
date: '`r format(Sys.time(), "%B %d, %Y")`'
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, output_file="/Users/miaforsline/Desktop/precip_comparisons/index.html") })
output:
  html_document:
    theme: flatly
    code_folding: hide
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 72
---

# Objectives

The goal of this step in the workflow is to compare S2 streamflow height
to precipitation data in order to quality check the streamflow data.

# Set Up

-   Install/load necessary R packages
-   Set working directory if necessary (or create a file path to use
    throughout the RMD to call the data from Box)

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
#install packages if necessary, then load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(
  here, #for file paths 
  janitor, 
  plotly, #for creating interactive plots 
  readxl, 
  tidyverse,
  zoo #for interpolating missing data points 
  )
```

-   Set constant values

```{r min_max_year}
#set constants 
min_year = 2017
max_year = 2021
```

-   Recreate intermediate data for later use

```{r intermediate_data, message=FALSE, warning=FALSE}
#2017 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#read in the 2017 data 
wy2017 <- read_excel(path = here(filepath, "WY2017.S2_Breakpoint.xlsx"), 
                     skip =3) #skip the top 3 rows of the Excel Sheet 

#clean the 2017 data
wy2017_clean <- wy2017 %>% 
  #change column names
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  #remove unnecessary columns 
  select(c("datetime", "stream_height_ft")) %>% 
  #format column names into lower_snake_case
  clean_names() %>% 
  #create a date column 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#2018

filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#create a function to specify custom time intervals 
`%between%` <- function(x, interval) x >= interval[1] & x <= interval[2]

#read in the 2018 data 
wy2018 <- read_excel(path = here(filepath, "WY2018.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2018 data
wy2018_clean <- wy2018 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#assign October NA values 
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-10-09', '2018-10-10'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-10-11', '2018-10-15'))] <- NA

#assign November NA values 
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-11-10', '2018-11-14'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-11-25', '2018-12-04'))] <- NA

#assign December NA values 
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-12-01', '2018-12-03'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-12-08', '2018-12-10'))] <- NA
wy2018_clean$stream_height_ft[wy2018_clean$date %between% as.Date(c('2018-12-11', '2018-12-31'))] <- NA

wy2018_altered <- wy2018_clean
wy2018_altered <- subset(wy2018_altered, !(wy2018_altered$date %between% as.Date(c('2018-10-31', '2018-11-01')) & wy2018_altered$stream_height_ft < 0.10))

filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#2019 
#read in the 2019 data 
wy2019 <- read_excel(path = here(filepath, "WY2019.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2019 data
wy2019_clean <- wy2019 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#assign February NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-01', '2019-04-02'))] <- NA

#assign March NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date == as.Date('2019-03-14')] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date == as.Date('2019-03-18')] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-03-30', '2019-03-31'))] <- NA

#assign April NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-01', '2019-04-02'))] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-08', '2019-04-17'))] <- NA
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-04-17', '2019-04-22'))] <- NA

#assign September NA values 
wy2019_clean$stream_height_ft[wy2019_clean$date %between% as.Date(c('2019-09-03', '2019-09-10'))] <- NA

#2020 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#read in the 2020 data 
wy2020 <- read_excel(path = here(filepath, "WY2020.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2020 data
wy2020_clean <- wy2020 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#create date range to fill in with NA values 
start_date <- as.POSIXct("7/28/2020 13:40:00", 
                         format = '%m/%d/%Y %H:%M:%S',
                         tz = "GMT")
end_date <- as.POSIXct("8/11/2020 00:00:00", 
                         format = '%m/%d/%Y %H:%M:%S',
                         tz = "GMT")

#create a function to specify custom time intervals 
`%between%` <- function(x, interval) x >= interval[1] & x <= interval[2]

#assign February NA values 
wy2020_clean$stream_height_ft[wy2020_clean$datetime %between% c(start_date, end_date)] <- NA

#2021 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/StripCharts/AnnualBreakPoint/S2"

#read in the 2021 data 
wy2021 <- read_excel(path = here(filepath, "WY2021.S2_Breakpoint.xlsx"), 
                     skip =3)

#clean the 2021 data
wy2021_clean <- wy2021 %>% 
  rename(datetime = "Date/time",
         stream_height_ft = "Stage.ft") %>% 
  select(c("datetime", "stream_height_ft")) %>% 
  clean_names() %>% 
  mutate(date = format(as.POSIXct(datetime, format = '%m/%d/%Y %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         date = as.POSIXct(date, tz = "GMT")) 

#identify date intervals that need to be filled w/NAs
date1 <- as.POSIXct("2021-03-04 00:00:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")
date2 <- as.POSIXct("2021-03-09 00:00:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")
date3 <- as.POSIXct("2021-11-11 23:59:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")
date4 <- as.POSIXct("2021-12-29 00:00:00", 
                         format = '%Y-%m-%d %H:%M:%S',
                         tz = "GMT")


#create a function to specify custom time intervals 
`%between%` <- function(x, interval) x >= interval[1] & x <= interval[2]

#assign March NA values
wy2021_clean$stream_height_ft[wy2021_clean$datetime %between% c(date1, date2)] <- NA

#assign Nov-Dec NA values 
wy2021_clean$stream_height_ft[wy2021_clean$datetime %between% c(date3, date4)] <- NA

#all water years 
all_streamflow <- rbind(wy2017_clean, 
                        wy2018_altered, 
                        wy2019_clean,
                        wy2020_clean,
                        wy2021_clean)

#air temp data
#set file path to read in data from Box 
filepath = "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/EDI/airT_RH_BP_30min/edi.990.1/data_objects/"

#read in the raw data 
air_temp_30 <- read_csv(here(filepath, "MEF_S2Bog_Met_30min.csv"))

#clean the data 
air_temp_30_clean <- air_temp_30 %>% 
  clean_names() %>% 
  rename(airtempc = air_t,
         rel_humidity = rh,
         datetime = timestamp) %>% 
  mutate(year = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year),
         airtempc_10 = (airtempc / 10),
         airtempc_100 = (airtempc / 100),
         #create a variable that indicates if the airtemp is above or below freezing
         freezing = 0
         ) %>% 
  pivot_longer(
    cols = c("airtempc", "airtempc_10", "airtempc_100"),
    names_to = "airtemp_type",
    values_to = "airtempC"
  )

#manual checks data 
filepath <- "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/Hydro/Streamflow/L0_subdaily/ManualChecks"

#read in the manual checks data 
mc <- read_csv(here(filepath, "2017-2021_S2Stage.csv"))

#clean the data
mc_clean <- mc %>% 
  clean_names() %>% 
  #remove S2 lagging pool data and keep only the S2 weir data
  subset(name == "S2 WEIR") %>%
  mutate(collected = as.POSIXct(collected, format = '%m/%d/%Y %H:%M', 
                                tz = "GMT"),
         year = format(as.POSIXct(collected, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         date = format(as.POSIXct(collected, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
         diff = point_gage - stripchart_stage
         ) %>% 
  subset(year >= min_year & year <= max_year) %>% 
  rename(datetime = collected)

#precipitation data 
#set file path to read in data from Box 
filepath = "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/EDI/precipitation_15min/edi.849.2/data_objects/"

#read in the raw data 
precip_15 <- read_csv(here(filepath, "MEF_South_precip_15min.csv"))

#clean the data 
precip_15_clean <- precip_15 %>% 
  clean_names() %>% 
  mutate(year = format(as.POSIXct(timestamp, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year)) %>% 
  subset(year >= min_year) %>% 
  rename(precip_in = south_pcp) %>% 
  mutate(precip_10 = (precip_in / 10)) %>% 
  pivot_longer(
    cols = c("precip_in", "precip_10"),
    names_to = "precip_type",
    values_to = "precip_in"
  ) %>% 
  rename(datetime = timestamp)

#read in 15min precip data 
precip_6h <- precip_15 

#clean and sum precip by 6 hours 
precip_6h_clean <- precip_6h %>% 
  clean_names() %>% 
  mutate(timestamp = lubridate::floor_date(timestamp, unit = "6 hours")) %>%
    group_by(timestamp) %>%
    dplyr::summarize(south_pcp = sum(south_pcp)) %>%
    ungroup() %>% 
  rename(datetime = timestamp,
         precip_in = south_pcp) %>% 
  mutate(
    year = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year),
     date = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
    precip_10_in = precip_in / 10
  ) %>% 
  subset(year >= min_year & year <= max_year)

#read in 15min precip data 
precip_12h <- precip_15 

#clean and sum precip by 6 hours 
precip_12h_clean <- precip_12h %>% 
  clean_names() %>% 
  mutate(timestamp = lubridate::floor_date(timestamp, unit = "12 hours")) %>%
    group_by(timestamp) %>%
    dplyr::summarize(south_pcp = sum(south_pcp)) %>%
    ungroup() %>% 
  rename(datetime = timestamp,
         precip_in = south_pcp) %>% 
  mutate(
    year = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year),
    date = format(as.POSIXct(datetime, format = '%Y-%m-%d %H:%M:%S', 
                                  tz = "GMT"), 
                       format = '%Y-%m-%d'),
    precip_10_in = precip_in / 10
  ) %>% 
  subset(year >= min_year & year <= max_year)

#set file path to read in data from Box 
filepath = "/Users/miaforsline/Library/CloudStorage/Box-Box/External-MEF_DATA/EDI/precipitation_daily/edi.563.3/data_objects/"

#read in the raw data 
precip_daily <- read_csv(here(filepath, "MEF_precipitation_daily.csv"))

#clean the data 
precip_daily_clean <- precip_daily %>% 
  clean_names() %>% 
  select(date, south_pcp, south_flag) %>% 
  mutate(year = format(as.POSIXct(date, format = '%Y-%m-%d', 
                                  tz = "GMT"), 
                       format = '%Y'),
         year = as.numeric(year)) %>% 
  subset(year >= min_year) %>% 
  rename(precip_in = south_pcp) %>% 
  mutate(precip_10 = (precip_in / 10)) %>% 
  pivot_longer(
    cols = c("precip_in", "precip_10"),
    names_to = "precip_type",
    values_to = "precip_in"
  )
```


# Streamflow compared to cumulative precipitation and manual checkpoints

For interactive plots,

-   cumulative precipitation is plotted as grey bars,

-   streamflow is plotted as a [blue line]{style="color:blue"},

-   and streamflow manual checkpoints are plotted as **black dots**.

# Cumulative Precipitation Every 6 Hours

```{r p_6h}
#calculate the ratio of the 2 y-axes to plot the variables together
trans_value <- max(precip_6h_clean$precip_10_in, na.rm = TRUE) / max(all_streamflow$stream_height_ft, na.rm = TRUE)

#plot 
p_6h <- ggplot() + 
  #precipitation data 
  geom_col(data = precip_6h_clean, 
            aes(x = datetime, 
                y = precip_10_in,
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Precipitation (in/10): ", precip_10_in)),
            colour = I("grey"),
            size = 0.25) + 
  #streamflow data 
  geom_line(data = all_streamflow,
             aes(x = datetime, 
                 y = stream_height_ft, 
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3))),
            color = "blue",
            size = 0.4) + 
  #manual check points
  geom_point(data = mc_clean, 
             aes(x = datetime, 
                 y = stripchart_stage, 
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height Checkpoint (ft): ", stripchart_stage)),
             size = 0.5) + 
  theme_classic() + 
  labs(x = "Time", 
       title = paste0("S2 Bog Streamflow and Cumulative \n Precipitation Every 6 Hours (", min_year, "-", max_year, ")"),
       subtitle = "Streamflow is plotted in blue. \n Precipitation is plotted in grey. \n Manual streamflow checks are plotted as black points.") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, 
                                     size = 7)) +
  scale_y_continuous(
    #features of the first axis
    name = "Stream Height (ft)",
    #add a second axis and specify its features
    ##determine the ratio between the max value of Axis 1 vs the max value of Axis 2 to decide what number to put in the trans argument
    sec.axis = sec_axis(trans = ~.*trans_value, name = "Cumulative Precipitation / 10 (in)")
  )

#static plot
#p_6h

#interactive plot 
ggplotly(p_6h, tooltip = "text") %>% 
  layout(title = list(text = paste0(paste0("S2 Bog Streamflow and Cumulative \n Precipitation Every 6 Hours (", min_year, "-", max_year, ")"),
                                    '<br>',
                                    '<sup>',
                                     paste0(" "), 
                                    '</sup>')), 
         font = list(size = 12)
         )

```

# Cumulative Precipitation Every 12 Hours

For interactive plots,

-   cumulative precipitation is plotted as grey bars,

-   streamflow is plotted as a [blue line]{style="color:blue"},

-   and streamflow manual checkpoints are plotted as **black dots**.

```{r p_12h}
#calculate the ratio of the 2 y-axes to plot the variables together
trans_value <- max(precip_12h_clean$precip_10_in, na.rm = TRUE) / max(all_streamflow$stream_height_ft, na.rm = TRUE)

#plot 
p_12h <- ggplot() + 
  #precipitation data 
  geom_col(data = precip_12h_clean, 
            aes(x = datetime, 
                y = precip_10_in, 
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Precipitation (in/10): ", precip_10_in)),
            colour = I("grey"),
            size = 0.25) + 
  #streamflow data 
  geom_line(data = all_streamflow,
             aes(x = datetime, 
                 y = stream_height_ft, 
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3))),
            color = "blue",
            size = 0.4) + 
  #manual check points
  geom_point(data = mc_clean, 
             aes(x = datetime, 
                 y = stripchart_stage, 
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height Checkpoint (ft): ", stripchart_stage)),
             size = 0.5) + 
  theme_classic() + 
  labs(x = "Time", 
       title = paste0("S2 Bog Streamflow and Cumulative \n Precipitation Every 12 Hours (", min_year, "-", max_year, ")"),
       subtitle = "Streamflow is plotted in blue. \n Precipitation is plotted in grey. \n Manual streamflow checks are plotted as black points.") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, 
                                     size = 7)) +
  scale_y_continuous(
    #features of the first axis
    name = "Stream Height (ft)",
    #add a second axis and specify its features
    ##determine the ratio between the max value of Axis 1 vs the max value of Axis 2 to decide what number to put in the trans argument
    ###for example, max value of precipitation / max value of streamflow = 0.533...
    sec.axis = sec_axis(trans = ~.*trans_value, name = "Cumulative Precipitation / 10 (in)")
  )

#static plot
#p_12h

#interactive plot 
ggplotly(p_12h, tooltip = "text")

```

# Cumulative Daily Precipitation

For interactive plots,

-   cumulative precipitation is plotted as grey bars,

-   streamflow is plotted as a [blue line]{style="color:blue"},

-   and streamflow manual checkpoints are plotted as **black dots**.

```{r p_daily}
#format precipitation data for a smooth joining process
precip_daily_join <- precip_daily_clean %>% 
  pivot_wider(
    names_from = "precip_type",
    values_from = "precip_in"
  ) %>% 
  subset(year >= min_year & year <= max_year) %>% 
  rename(precip_10_in = precip_10) %>% 
  mutate(datetime = format(as.POSIXct(date, format = '%Y-%m-%d', 
                                  tz = "GMT"),
                           format = '%Y-%m-%d %H:%M:%S'),
         datetime = as.POSIXct(datetime))

#join streamflow and daily precipitation data 
streamflow_precip_daily <- full_join(x = all_streamflow, 
                                     y = precip_daily_join, 
                                     by = "date") %>% 
  select(-precip_in) %>% 
  subset(year >= min_year & year <= max_year)

#calculate the ratio of the 2 y-axes to plot the variables together
trans_value <- max(precip_daily_join$precip_10_in, na.rm = TRUE) / max(all_streamflow$stream_height_ft, na.rm = TRUE)

#plot 
p_daily <- ggplot() + 
  #precipitation data 
  geom_col(data = precip_daily_join, 
            aes(x = datetime, 
                y = precip_10_in,
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "Precipitation (in/10): ", precip_10_in)),
            colour = I("grey"),
            size = 0.25) + 
  #streamflow data 
  geom_line(data = all_streamflow,
             aes(x = datetime, 
                 y = stream_height_ft, 
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3))),
            color = "blue",
            size = 0.4) + 
  #manual check points
  geom_point(data = mc_clean, 
             aes(x = datetime, 
                 y = stripchart_stage, 
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height Checkpoint (ft): ", stripchart_stage)),
             size = 0.5) + 
  theme_classic() + 
  labs(x = "Time", 
       title = paste0("S2 Bog Streamflow and Cumulative \n Daily Precipitation (", min_year, "-", max_year, ")"),
       subtitle = "Streamflow is plotted in blue. \n Precipitation is plotted in grey. \n Manual streamflow checks are plotted as black points.") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, 
                                     size = 7)) +
  scale_y_continuous(
    #features of the first axis
    name = "Stream Height (ft)",
    #add a second axis and specify its features
    ##determine the ratio between the max value of Axis 1 vs the max value of Axis 2 to decide what number to put in the trans argument
    ###for example, max value of precipitation / max value of streamflow = 0.533...
    sec.axis = sec_axis(trans = ~.*trans_value, name = "Cumulative Daily Precipitation / 10 (in)")
  )

#static plot
#p_daily

#interactive plot 
ggplotly(p_daily, tooltip = "text")
```

# Code Accessibility

Note that this file is being knit as
[`index.html`](https://github.com/miaforsline/precip_comparisons/blob/main/index.html)
into the public [`precip_comparisons` GitHub
repository](https://github.com/miaforsline/precip_comparisons) in order
to update the [GitHub pages
website](https://miaforsline.github.io/precip_comparisons/), where the
plots can be viewed online.
