---
title: 'Marcell Experimental Forest: S2 Stream Height & Air Temperature Comparison Analyses'
author: "Mia Forsline"
date: '`r format(Sys.time(), "%B %d, %Y")`'
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, output_file="/Users/miaforsline/Desktop/air_comparisons/index.html") })
output:
  html_document:
    theme: flatly
    code_folding: hide
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 72
---

# Set Up

-   Install/load necessary R packages
-   Set working directory if necessary (or create a file path to use
    throughout the RMD to call the data from Box)

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
#install packages if necessary, then load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(
  here, #for file paths 
  janitor, 
  plotly, #for creating interactive plots 
  readxl, 
  tidyverse,
  zoo #for interpolating missing data points 
  )
```

Read in data from CSVs in the `intermediate_data` directory

```{r csv, message=FALSE, warning=FALSE}
#read in necessary CSV files from prior RMDs
all_streamflow <- read_csv(here("intermediate_data", "all_streamflow.csv"))
air_temp_30_clean <- read_csv(here("intermediate_data", "air_temp_30_clean.csv"))
mc_clean <- read_csv(here("intermediate_data", "mc_clean.csv")) %>% 
  rename(datetime = collected)
precip_15_clean <- read_csv(here("intermediate_data", "precip_15_clean.csv")) 
precip_6h_clean <- read_csv(here("intermediate_data", "precip_6h_clean.csv")) 
precip_12h_clean <- read_csv(here("intermediate_data", "precip_12h_clean.csv")) 
precip_daily_clean <- read_csv(here("intermediate_data", "precip_daily_clean.csv")) 
precip_daily_join <- read_csv(here("intermediate_data", "precip_daily_join.csv")) 
```

Set constant values (aka the date range of the analysis)

```{r min_max_year}
#set constants 
min_year = 2017
max_year = 2021
```

Create sub-directories if necessary

```{r directories}
output_dir <- file.path(here("intermediate_data"))

if (!dir.exists(output_dir)){
dir.create(output_dir)
} else {
    print("intermediate_data directory already exists!")
}

output_dir <- file.path(here("figures"))

if (!dir.exists(output_dir)){
dir.create(output_dir)
} else {
    print("figures directory already exists!")
}
```

# Daily Cumulative Precipitation + Air Temperature Every 30min

For interactive plots, cumulative precipitation is plotted as grey bars,
and air temperature is plotted as a red line.

```{r p_daily}
#format precipitation data for plotting 
precip_15_join <- precip_15_clean %>% 
  pivot_wider(
    names_from = "precip_type",
    values_from = "precip_in"
  ) %>% 
  subset(year >= min_year & year <= max_year) %>% 
  rename(precip_10_in = precip_10)

#format air temperature data for plotting 
airtemp_join <- air_temp_30_clean %>% 
  pivot_wider(
    names_from = "airtemp_type",
    values_from = "airtempC"
  ) %>% 
  subset(year >= min_year & year <= max_year) %>% 
  mutate(freezing = as.factor(freezing))

#format air temp data for plotting 
precip_air <- full_join(x = precip_15_join, 
                        y = airtemp_join, 
                        by = c("datetime", "year")) %>% 
  subset(year >= min_year & year <= max_year)

#join air and precipitation data 
precip_air <- full_join(x = precip_15_join, 
                        y = airtemp_join, 
                        by = c("datetime", "year")) %>% 
  subset(year >= min_year & year <= max_year)

#plot
p_daily <- ggplot() + 
  #airtemp data
  geom_line(data = airtemp_join, 
            aes(x = datetime, 
                y = airtempc_100,
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "AirTemp (ºC/100): ", airtempc_100)
                ),
            color = "pink",
            alpha = 0.5,
            size = 0.25) + 
  #precipitation data 
  geom_col(data = precip_daily_join, 
            aes(x = datetime, 
                y = precip_10_in,
                colour = I("grey"),
                group = 1,
                text = paste0("Date: ", date, "\n",
                              "Daily Cumulative Precipitation (in/10): ", precip_10_in)),
            size = 0.25) +
  theme_classic() + 
  labs(x = "Time", 
       y = " ", 
       title = "S2 Daily Cumulative Precipitation \n and Air Temperature Every 30min",
       subtitle = "Daily cumulative precipitation is plotted in grey. \n Air temperature is plotted in pink.") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, 
                                     size = 7)) 

# + 
#   scale_y_continuous(
#     # Features of the first axis
#     name = "Air Temperature (ºC) / 100",
#     # Add a second axis and specify its features
#     sec.axis = sec_axis(trans = ~., name = "Cumulative Precipitation (in) / 10")
#   ) 

#static plot
#p_daily

# #save figure 
# ggsave(filename = "precip_daily_airtemp_30.png",
#        plot = p_daily,
#        path = "figures/",
#        width = 10,
#        height = 5,
#        units = c("in"),
#        dpi = 300)

#interactive plot
ggplotly(p_daily,
         tooltip = "text")
```

# Streamflow + air temperature data (every 30min) + manual checks

```{r p_air}
#calculate the ratio of the 2 y-axes to plot the variables together
trans_value <- max(airtemp_join$airtempc_100, na.rm = TRUE) / max(all_streamflow$stream_height_ft, na.rm = TRUE)

#plot 
p_air <- ggplot() + 
  #airtemp data
  geom_line(data = airtemp_join, 
            aes(x = datetime,
                y = airtempc_100, 
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "AirTemp (ºC/100): ", airtempc_100)
                ),
            color = "pink", 
            alpha = 0.4) +
  #streamflow data 
  geom_line(data = all_streamflow,
             aes(x = datetime, 
                 y = stream_height_ft,
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3))
                 ),
            color = "blue",
            size = 0.75) + 
  #manual check points
  geom_point(data = mc_clean, 
             aes(x = datetime, 
                 y = stripchart_stage,
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", datetime, "\n",
                              "Manual Stream Height Check (ft): ", stripchart_stage)
                 ),
             size = 0.5) + 
  theme_classic() + 
  labs(x = "Time", 
       y = " ",
       title = paste0("S2 Bog Streamflow and Air Temperature (", min_year, "-", max_year, ")"),
       subtitle = "Air temperature is plotted in pink. \n Streamflow is plotted in blue. \n Manual streamflow checkpoints are plotted as black dots.") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, 
                                     size = 7)) + 
  scale_y_continuous(
    # Features of the first axis
    name = "Air Temperature / 100 (ºC)",
    # Add a second axis and specify its features
    sec.axis = sec_axis(trans = ~.*trans_value, name = "Stream Height (ft)")
  ) 

#static plot
#p_air

# #save the plot in the figures folder 
# ggsave(filename = "streamflow_airtemp_mc.png",
#        plot = p_air,
#        path = "figures/",
#        width = 8,
#        height = 4,
#        units = c("in"),
#        dpi = 300)

#interactive plot
ggplotly(p_air,
         tooltip = "text")
```

# Streamflow + precip (daily) + air temp + manual checks

```{r p_all}
#plot 
p_all <- ggplot() + 
  #airtemp data
  geom_line(data = airtemp_join, 
            aes(x = datetime,
                y = airtempc_100, 
                group = 1,
                #create hovering labels for interactive graph 
                text = paste0("DateTime: ", datetime, "\n",
                              "AirTemp (ºC/100): ", airtempc_100)
                ),
            color = "pink", 
            alpha = 0.4) +
  #streamflow data 
  geom_line(data = all_streamflow,
             aes(x = datetime, 
                 y = stream_height_ft,
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", datetime, "\n",
                              "Stream Height (ft): ", round(stream_height_ft, digits = 3))
                 ),
            color = "blue",
            size = 0.75) + 
  #manual check points
  geom_point(data = mc_clean, 
             aes(x = datetime, 
                 y = stripchart_stage,
                 group = 1,
                 #create hovering labels for interactive graph 
                 text = paste0("DateTime: ", datetime, "\n",
                              "Manual Stream Height Check (ft): ", stripchart_stage)
                 ),
             size = 0.5) + 
  #daily precipitation data 
  geom_col(data = precip_daily_join, 
            aes(x = datetime, 
                y = precip_10_in,
                colour = I("grey"),
                group = 1,
                text = paste0("Date: ", date, "\n",
                              "Daily Cumulative Precipitation (in/10): ", precip_10_in)),
            size = 0.25) +
  theme_classic() + 
  labs(x = "Time", 
       y = " ", 
       title = paste0("S2 Bog Streamflow, Air Temperature, \n and Daily Cumulative Precipitation (", min_year, "-", max_year, ")"),
       subtitle = "Air temperature is plotted in pink and scaled by 100. 
       Precipitation is plotted as grey bars and scaled by 10.
       Streamflow is plotted in blue.
       Manual streamflow checkpoints are plotted as black dots.") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, 
                                     size = 7)) 
# + 
#   scale_y_continuous(
#     # Features of the first axis
#     name = "Air Temperature / 100 (ºC)",
#     # Add a second axis and specify its features
#     sec.axis = sec_axis(trans = ~.*trans_value, name = "Stream Height (ft)")
#   ) 

#static plot
#p_all

# #save the plot in the figures folder
# ggsave(filename = "airtemp_mc_precip_streamflow.png",
#        plot = p_all,
#        path = "figures/",
#        width = 8,
#        height = 4,
#        units = c("in"),
#        dpi = 300)

#interactive plot
ggplotly(p_all,
         tooltip = "text")
```

Note that this file is being knit as
[`index.html`](https://github.com/miaforsline/air_comparisons/blob/main/index.html)
into the [`air_comparisons`
repository](https://github.com/miaforsline/precip_comparisons) in order
to update the [GitHub pages
website](https://miaforsline.github.io/air_comparisons/), where the
plots can be viewed online.
